A significance test in forward stepwise model selection

### 2014-05-08 paper/paper.tex
Small example, note on power, various edits

\begin{abstract}
% perhaps de-emphasize group LASSO and emphasize forward stepwise?
  We extend the methods developed by \cite{significance:LASSO} and
  \cite{tests:adaptive} on significance tests for penalized
  regression to stepwise model selection by iteratively applying
  the global null hypothesis test on the residual in a forward stepwise
  procedure. While not fully theoretically justified, the resulting method
  has the computational strengths of stepwise selection and partially
  solves the problem of inflated test statistics due to model selection.
  We illustrate the flexibility of this method by applying it to
  novel specialized applications of forward stepwise to a hierarchically
  constrained interactions model and generalized additive models.
\end{abstract}

@@

In the LASSO setting, \cite{significance:LASSO} derived
a novel test statistic and its asymptotic null distribution, making
possible valid inferences after model selection using the full data.
\cite{tests:adaptive} modified and extended those results to the
group LASSO \citep{groupLASSO} and other penalized regression
problems, but only under the global null hypothesis.
One of the strengths of these test statistics is that they can be
used for valid significance testing when computed on the same
data used for model selection, eliminating the need for data splitting
or for situations in which data splitting is not appropriate. For example,
when categorical covariates contain rare levels with very few observations
it can be difficult or impossible to split the sample with an adequate
number of such observations occurring in each split.

@@

In the next section we establish notation and describe the
forward stepwise algorithm used in most of the paper.
Section~\ref{sec:testing} reviews some recent work on
post-selection inference
\citep{significance:LASSO,tests:adaptive,LASSO:fixed}
relevant to our work here, and describes a general framework
for post-selection inference based on quadratic comparisons.
Simulation results in Section~\ref{sec:simulations} show
empirically that our method performs well in settings where forward
stepwise itself performs well, and that
various stopping rules using our test statistic--including
some from \cite{sequential:fdr}--appear promising. In
Section~\ref{sec:applications} we apply
the method to several variants of forward
stepwise tailored to models with interactions and generalized additive
models, as well as to a real data example.
(\todo Change this paragraph if necessary as the sections are completed)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Performance of forward stepwise}

@@

So far we have left open the question of choosing among the models in
the forward stepwise sequence, i.e. when to stop stepping
forward. Some approaches for this problem can be posed as optimization
criteria which stop at the step minimizing
\begin{equation}
\begin{aligned}
\label{eq:subsetregress}
\frac{1}{2} \| y - X \beta_s \|_2^2 + \lambda \pen(\beta_s)
\end{aligned}
\end{equation}
where we have written $\{ \beta_s : s = 1, \ldots, steps \}$ as
the sequence of models output by forward stepwise. The function
$\pen(\beta)$ is a penalty on model complexity usually taken to be the
number of nonzero entries of $\beta$. Proposals for $\lambda$ include
2 ($C_p$ of \cite{CP}, AIC of \cite{AIC}), $\log(n)$ (BIC of \cite{BIC}), and
$2\log(p)$ (RIC of \cite{RIC}). Stopping rules based on classical test
statistics have also been used, so it is natural to consider using
the new test statistics of \cite{significance:LASSO} or
\cite{tests:adaptive}
to choose a model. \cite{sequential:fdr} examined some stopping rules
using the asymptotic p-values of \cite{significance:LASSO} and showed
their stopping rules control false discovery rate--the expected
proportion of noise variables among variables declared significant
\citep{fdr}. We explore this further in Section~\ref{sec:simulations}.

@@

For various sparsity levels $k$, we applied Algorithm~\ref{algo:fs} to
various data sets with \textit{steps} set to $k$.
In the resulting active set $A_k$, the number of true variables divided by $k$ is the $k$-True Positive Proportion ($k$-TPP):
\[
\text{TPP}_k = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
\]
In some plots this has been denoted $k$-SP instead of $k$-TPP (\todo fix).
Since we know $k$ in the simulations we can use $k$-TPP as a measure of how well forward stepwise is performing. When it is close to 1 we are recovering most of the true variables before including false ones.
In our simulations nonzero coefficients have magnitudes
in a range of multiples of $\gamma := \sqrt{2\log(G)/n}$, e.g. in $[1.4
\gamma, 1.6 \gamma]$. Results are shown in
Figure~\ref{fig:fwdstepsim}. After computing the
coherence of these design matrices, some calculations show the required
sparsity level to guarantee exact recovery in these situations is
about 2 or smaller, and the required nonzero coefficient magnitude is
likely in the range of 10 to 100 times $\gamma$. These simulations are far
from the stringent conditions required by theory to guarantee perfect
recovery, but the performance, while not perfect, may still be good
enough for some purposes. Finally, to serve as a comparison with existing
commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}(b)
we show in green the TPP and median model sizes for models chosen by
the \textit{step} function in \textit{R} using the BIC criterion, and
the same for RIC are plotted in purple.

@@

\section{Significance testing: from LASSO to group LASSO}
\label{sec:testing}

In the ordinary least squares setting, a significance test for a
single variable can be conducted by comparing the drop in residual
sums of squares (RSS) to a $\chi^2_1$ distribution. Similarly, when
adding a group of $k$ variables we can compare the drop in RSS to a
$\chi^2_k$ random variable. This generally does not work when the
group to be added has been chosen by a method that uses the data
\citep{olshen:flevel},
and in particular it fails for forward stepwise procedures adding
the ``best'' (e.g. most highly correlated) predictor in each step. In
that case, the drop in RSS does not match its
theoretical null distribution even when the null hypothesis is
true. \cite{significance:LASSO} introduced a new test statistic based
on the knots in the LASSO solution path. They derived a simple
asymptotic null distribution, proved a type of convergence under broad
``minimum growth'' conditions, and demonstrated in simulations that
the test statistic closely matches its asymptotic distribution even in
finite samples.  That work marked an important advance in the problem
of combining inference with model selection.  \cite{tests:adaptive}
extended that work to the group LASSO \citep{groupLASSO} and other
problems, and modified the test statistic to one with an exact finite
sample distribution under the global null hypothesis.

To describe the previous work we require some facts about the solution path
of the LASSO.
Writing $\hat \beta(\lambda)$ for the LASSO solution for a fixed value of $\lambda$, we recall
the following facts summarized in \cite{significance:LASSO,tibshirani_LASSO_uniqueness}.

@@

The covariance test (assuming groups of size one and $\Sigma = \sigma^2 I$ is given at the first step by
\begin{equation}
\label{eq:covtest}
T_1 = \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \overset{n,p \to \infty}{\to} \text{Exp}(1).
\end{equation}
In \cite{tests:adaptive} it is pointed out 
\begin{equation}
\label{test:exact}
\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}(0,1).
\end{equation}
Further, the above test statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
{\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
The limiting results about later steps in \cite{significance:LASSO} can
%(formally)
be interpreted as
recursively applying $T_1$ to the variables that had not previously been selected by LARS.

@@

Recall the LASSO estimator is given by
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
\label{eq:LASSO}
\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
   \lambda \| \beta \|_1 \\
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \emph{group LASSO estimator}, a generalization of the LASSO,
 is a solution to the following penalized
least-squares convex problem
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
\label{eq:gsoln}
\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
   \lambda {\cal P}(\beta) \\
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
with the group penalty
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \begin{aligned}
  \label{eq:gpen}
    {\cal P}(\beta)&= \sum_{g=1}^G w_g \| \beta_g \|_2 \\
  \end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The parameter $\lambda \geq 0$ enforces sparsity in groups: for large
$\lambda$ most of the $\beta_g$ will be zero vectors. The weights
$w_g$ are usually taken to be $\sqrt {p_g}$ to normalize the penalty
across groups with different sizes.  This can also be accomplished
by scaling the corresponding submatrices by their Frobenius norms
and setting all weights equal.
Note that this includes the usual LASSO estimator as a
special case when all groups are of size 1, since then the
penalty term is the $L^1$-norm of $\beta$. The solution path of the group
LASSO has similar properties to the LASSO case, however it is not
generally piecewise linear.

For sufficiently large $\lambda$ the solution to~\eqref{eq:gsoln} is forced
to be zero. The smallest such $\lambda$ is denoted
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \begin{aligned}
   \lambda_1 &= \inf \{ \lambda \geq 0 : \hat \beta_{\lambda'} = 0 \text{ for all } \lambda' > \lambda \} \\
  \end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the group LASSO this is explicitly computable by
\begin{equation}
\label{eq:lammax}
\lambda_1^{\text{group}} = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2 . 
\end{equation}
The value above is also the dual norm of the penalty~\eqref{eq:gpen}
hence it can be expressed as
\begin{equation}
\label{eq:lammax}
\lambda_1^{\text{group}}
% = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2
 = \frac{1}{w_{g^*}}\eta_{g^*}^TX_{g^*}^Ty . 
\end{equation}
where $g^*$ is the group that achieves the maximum and 
$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the unit vector that achieves the norm $\|X_{g^*}^Ty\|_2$. One of the key conceptual points about our
test is that it conditions on both the maximizer $g^*$ and on the unit
vector $\eta_{g^*}$. Most of the p-value computations will be done in
terms of these quantities.

@@

% \begin{figure}
% \label{fig:polytope}
% \begin{center}
% \resizebox{!}{3in}{\input{../figs/polytope.pdf_tex}}
% \end{center}
% A polytope selection event. Solving the LASSO \eqref{eq:LASSO} for a fixed value
% of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
% active variables $z_{\hat{E}(y)}$. The event $\{y: (\hat{E}(y), z_{\hat{E}(y)}) = (E, z_E)\}$
% can be written as a set of affine constraints.
% \end{figure}

\subsection{Selection events}

Before giving the full description of the p-value, we turn to another
previous work and describe how to extend the framework there to the
setting with groups.
The approach to post-selection inference in \cite{LASSO:fixed}
describes the selection event for the LASSO as a convex polytope
satisfying a list of affine constraints.
%stylized in Figure \ref{fig:polytope}.
If all groups are of size 1, then
the event that we observe a given set of variables
in the forward stepwise path (with their signs as they enter) would similarly
be given by a set of affine inequalities.
%This is described in more detail in \cite{exact:lars}.
After selection, exact inference for a (randomly) chosen contrast $\eta(y)^T\mu$ of the mean vector
$\mu$ could then be accomplished by analyzing the conditional distribution
\begin{equation}
y \sim N(\mu, \sigma^2 I) \bigl| Ay \leq b.
\end{equation}
See \cite{LASSO:fixed,exact:lars} for further details on this approach.

With non-trivial groups, however, the event that $g^*(y)$, the first
group chosen, is equal to some fixed group $g$ is generally not given by
a set of affine constraints. Rather, considering line 3 of Algorithm~\ref{algo:fs}, we see
\begin{equation}
\label{eq:first:quadratic}
\begin{aligned}
\{g^*(y)=g\} &= \{ \|X_g^Ty\|_2 /w_g \geq \|X_h^Ty\|_2 /w_h , \ \forall h \neq g\} \\
&=  \{ y^T(X_g^TX_g)y / w_g^2 - y^T(X_h^TX_h)y / w_h^2 \geq 0, \ \forall h \neq g\} \\
\end{aligned}
\end{equation}
Hence, our selection event can be expressed as 
the intersection of a list of quadratic inequalities. This non-affine selection event is stylized in Figure
\ref{fig:curved}.
In the rest of this section, we consider an arbitrary selection procedure
$S$ that returns one of a set of possible outcomes $s \in {\cal S}$ by a set of quadratic inequalities.
That is, for all possible outcomes $s \in {\cal S}$ there is a set of indices
$I(s)$ such that
\begin{equation}
\label{eq:selection}
\begin{aligned}
\{S(y)=s\} = \cap_{i \in I(s)} \{y: y^TQ_iy + a_i^Ty\leq b_i \}.
\end{aligned}
\end{equation}
The quadratic forms are not assumed non-negative definite, but we can, without loss of generality
assume they are symmetric.
As in \cite{LASSO:fixed} we will be interested in some randomly chosen contrast
$\eta(S(y))^T\mu$.
In our grouped selection procedure, the selection rule is $S(y)=g^*(y)$
and $\eta$ is typically 
$$
\eta(S(y)) = \eta(g^*(y)) = \frac{X_{g^*(y)}^Ty}{\|X_{g^*(y)}^Ty\|_2}.
$$

@@

%\subsection{Relation to covariance test}
%
%Consider the simple case for the first covariate to enter the model.
%In this case the covariance test statistic is $T = \lambda_1(\lambda_1 - M)$.
%\todo Does $M = \lambda_2$? Check proof, now that statistic has changed
%
%This is Lemma 5 from \cite{significance:LASSO}.
%
%\begin{itemize}
%
%  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
%  \item Convergence to the limiting Exp(1) distribution is too slow
%    \[
%      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
%    \]
%    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
%  \item The limiting distribution only depends on $T$, but we also observe $M$
%  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
%  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
%    \[
%      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
%    \]
%    we can just use the left hand side
%  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
%  \item Now our p-value is
%    \[
%      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
%    \]   
%\end{itemize}
%
%

@@

We performed simulations with several classes of design matrices including
Gaussian, categorical, and categorical from a real data set. Gaussian
design matrices are generated either independently or with some global
correlation $\rho > 0$ between all pairs of columns. Categorical matrices
were generated by first choosing a vector of probabilities for a given
variable from a Dirichlet prior, and then sampling category levels with
that vector of probabilities. Resampling was used to ensure the minimum
number of observations in any level was at least five. Finally, categorical
varables were encoded as groups of 0-1 vectors using the full encoding.
The categorical data matrices from real data are described in
Section~\ref{sec:hiv}

>>>>

\begin{abstract}
% perhaps de-emphasize group LASSO and emphasize forward stepwise?
  We extend the methods developed by \cite{significance:lasso} and
  \cite{tests:adaptive} on significance tests for penalized
  regression to stepwise model selection by iteratively applying
  the global null hypothesis test on the residual in a forward stepwise
  procedure. While not fully theoretically justified, the resulting method
  has the computational strengths of stepwise selection and partially
  solves the problem of inflated test statistics due to model selection.
  We illustrate the flexibility of this method by applying it to
  novel specialized applications of forward stepwise to a hierarchically
  constrained interactions model and generalized additive models.
\end{abstract}

@@

In the LASSO setting, \cite{significance:lasso} derived
a novel test statistic and its asymptotic null distribution, making
possible valid inferences after model selection using the full data.
\cite{tests:adaptive} modified and extended those results to the
group LASSO \citep{grouplasso} and other penalized regression
problems, but only under the global null hypothesis.
One of the strengths of these test statistics is that they can be
used for valid significance testing when computed on the same
data used for model selection, eliminating the need for data splitting
or for situations in which data splitting is not appropriate. For example,
when categorical covariates contain rare levels with very few observations
it can be difficult or impossible to split the sample with an adequate
number of such observations occurring in each split.

@@

As a small illustrative example, consider a response $Y$ with ten
categorical predictors $X_g$ having between 2 and 4 levels. The true 
relationship is that $Y$ only depends on $X_1$ and $X_9$. Specifically,
if observation $i$ has covariates $X_{1,i} = j, X_{9,i} = k$ then
\[
    Y_i = \beta_{1,j} + \beta_{9,k} + \epsilon_i , \qquad
    \epsilon_i \sim N(0,1)
\]
$X_1$ has three levels with $\beta_1 = (1, .5, -1)$, $X_9$
has two levels with $\beta_9 = (.5, -.5)$. We generated a random
categorical design matrix $X$ and created an instance of this example
with $n = 40$ observations. Running forward stepwise, we calculated
both our p-value and the usual $\chi^2$ p-value at each step. In
Table~\ref{tab:example} we see that forward stepwise chooses the
truly nonzero variables first and both p-values for these are small.
However, once forward stepwise begins adding noise variables, the
$\chi^2$ p-value remains relatively small, potentially leading to 
incorrect inferences.

\input{../tables/small.tex}

In the next section we establish general notation and describe the
forward stepwise algorithm used throughout the paper.
Section~\ref{sec:testing} reviews some recent work on
post-selection inference
\citep{significance:lasso,tests:adaptive,lasso:fixed}
relevant to our work here, and describes a general framework
for post-selection inference based on quadratic comparisons.
Simulation results in Section~\ref{sec:simulations} show
empirically that our method performs well in settings where forward
stepwise itself performs well, and that
various stopping rules using our test statistic--including
some from \cite{sequential:fdr}--appear promising. In
Section~\ref{sec:applications} we apply
the method to several variants of forward
stepwise tailored to models with interactions and generalized additive
models, as well as to a real data example involving genomic prediction
of individual drug responses to treat HIV.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Performance of forward stepwise}
\label{sec:fs:performance}

@@

So far we have left open the question of choosing among the models in
the forward stepwise sequence, i.e. when to stop stepping
forward. Some approaches for this problem can be posed as optimization
criteria which stop at the step minimizing
\begin{equation}
\begin{aligned}
\label{eq:subsetregress}
\frac{1}{2} \| y - X \beta_s \|_2^2 + \lambda \pen(\beta_s)
\end{aligned}
\end{equation}
where we have written $\{ \beta_s : s = 1, \ldots, steps \}$ as
the sequence of models output by forward stepwise. The function
$\pen(\beta)$ is a penalty on model complexity usually taken to be the
number of nonzero entries of $\beta$. Proposals for $\lambda$ include
2 ($C_p$ of \cite{CP}, AIC of \cite{AIC}), $\log(n)$ (BIC of \cite{BIC}), and
$2\log(p)$ (RIC of \cite{RIC}). Stopping rules based on classical test
statistics have also been used, so it is natural to consider using
the new test statistics of \cite{significance:lasso} or
\cite{tests:adaptive}
to choose a model. \cite{sequential:fdr} examined some stopping rules
using the asymptotic p-values of \cite{significance:lasso} and showed
their stopping rules control false discovery rate--the expected
proportion of noise variables among variables declared significant
\citep{fdr}. We explore this further in Section~\ref{sec:simulations}.

@@

For various sparsity levels $k$, we applied Algorithm~\ref{algo:fs} to
various data sets with \textit{steps} set to $k$.
In the resulting active set $A_k$, the number of true variables divided by $k$ is the $k$-True Positive Proportion ($k$-TPP):
\[
k\text{-TPP} = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
\]
Since we know $k$ in the simulations we can use $k$-TPP as a measure of how well forward stepwise is performing. When it is close to 1 we are recovering most of the true variables before including false ones.
In our simulations nonzero coefficients have magnitudes
in a range of multiples of $\gamma := \sqrt{2\log(G)/n}$, e.g. in $[1.4
\gamma, 1.6 \gamma]$. Results are shown in
Figure~\ref{fig:fwdstepsim}. After computing the
coherence of these design matrices, some calculations show the required
sparsity level to guarantee exact recovery in these situations is
about 2 or smaller, and the required nonzero coefficient magnitude is
likely in the range of 10 to 100 times $\gamma$. These simulations are far
from the stringent conditions required by theory to guarantee perfect
recovery, but the performance, while not perfect, may still be good
enough for some purposes. Finally, to serve as a comparison with existing
commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}(b)
we show in green the TPP and median model sizes for models chosen by
the \textit{step} function in \textit{R} using the BIC criterion, and
the same for RIC are plotted in purple.

@@

\section{Significance testing with model selection}
\label{sec:testing}

In the ordinary least squares setting, a significance test for a
single variable can be conducted by comparing the drop in residual
sums of squares (RSS) to a $\chi^2_1$ distribution. Similarly, when
adding a group of $k$ variables we can compare the drop in RSS to a
$\chi^2_k$ random variable. This generally does not work when the
group to be added has been chosen by a method that uses the data
\citep{olshen:flevel},
and in particular it fails for forward stepwise procedures adding
the ``best'' (e.g. most highly correlated) predictor in each step. In
that case, the drop in RSS does not match its
theoretical null distribution even when the null hypothesis is
true. \cite{significance:lasso} introduced a new test statistic based
on the knots in the LASSO solution path. They derived a simple
asymptotic null distribution, proved a type of convergence under broad
``minimum growth'' conditions, and demonstrated in simulations that
the test statistic closely matches its asymptotic distribution even in
finite samples.  That work marked an important advance in the problem
of combining inference with model selection.  \cite{tests:adaptive}
extended that work to the group LASSO \citep{grouplasso} and other
problems, and modified the test statistic to one with an exact finite
sample distribution under the global null hypothesis.

To describe the previous work we require some facts about the solution path
of the LASSO.
Writing $\hat \beta(\lambda)$ for the LASSO solution for a fixed value of $\lambda$, we recall
the following facts summarized in \cite{significance:lasso,tibshirani_LASSO_uniqueness}.

@@

The covariance test (assuming groups of size one and $\Sigma = \sigma^2 I$ is given at the first step by
\begin{equation}
\label{eq:covtest}
T_1 = \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \overset{n,p \to \infty}{\to} \text{Exp}(1).
\end{equation}
In \cite{tests:adaptive} it is pointed out 
\begin{equation}
\label{test:exact}
\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}(0,1).
\end{equation}
Further, the above test statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
{\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
The limiting results about later steps in \cite{significance:lasso} can
%(formally)
be interpreted as
recursively applying $T_1$ to the variables that had not previously been selected by LARS.

@@

Recall the LASSO estimator is given by
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
\label{eq:lasso}
\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
   \lambda \| \beta \|_1 \\
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \emph{group LASSO estimator}, a generalization of the LASSO,
 is a solution to the following penalized
least-squares convex problem
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
\label{eq:gsoln}
\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
   \lambda {\cal P}(\beta) \\
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
with the group penalty
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \begin{aligned}
  \label{eq:gpen}
    {\cal P}(\beta)&= \sum_{g=1}^G w_g \| \beta_g \|_2 \\
  \end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The parameter $\lambda \geq 0$ enforces sparsity in groups: for large
$\lambda$ most of the $\beta_g$ will be zero vectors. The weights
$w_g$ are usually taken to be $\sqrt {p_g}$ to normalize the penalty
across groups with different sizes.  This can also be accomplished
by scaling the corresponding submatrices by their Frobenius norms
and setting all weights equal.
Note that this includes the usual LASSO estimator as a
special case when all groups are of size 1, since then the
penalty term is the $L^1$-norm of $\beta$. The solution path of the group
LASSO has similar properties to the LASSO case, however it is not
generally piecewise linear.

For sufficiently large $\lambda$ the solution to~\eqref{eq:gsoln} is forced
to be zero. The smallest such $\lambda$ is denoted
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \begin{aligned}
   \lambda_1 &= \inf \{ \lambda \geq 0 : \hat \beta_{\lambda'} = 0 \text{ for all } \lambda' > \lambda \} \\
  \end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the group LASSO this is explicitly computable by
\begin{equation}
%\label{eq:lammax}
\lambda_1^{\text{group}} = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2 . 
\end{equation}
The value above is also the dual norm of the penalty~\eqref{eq:gpen}
hence it can be expressed as
\begin{equation}
\label{eq:lammax}
\lambda_1^{\text{group}}
% = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2
 = \frac{1}{w_{g^*}}\eta_{g^*}^TX_{g^*}^Ty . 
\end{equation}
where $g^*$ is the group that achieves the maximum and 
$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the unit vector that achieves the norm $\|X_{g^*}^Ty\|_2$. One of the key conceptual points about our
test is that it conditions on both the maximizer $g^*$ and on the unit
vector $\eta_{g^*}$. Most of the p-value computations will be done in
terms of these quantities.

@@

% \begin{figure}
% \label{fig:polytope}
% \begin{center}
% \resizebox{!}{3in}{\input{../figs/polytope.pdf_tex}}
% \end{center}
% A polytope selection event. Solving the LASSO \eqref{eq:lasso} for a fixed value
% of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
% active variables $z_{\hat{E}(y)}$. The event $\{y: (\hat{E}(y), z_{\hat{E}(y)}) = (E, z_E)\}$
% can be written as a set of affine constraints.
% \end{figure}

\subsection{Quadratic selection events}

Before giving the full description of the p-value, we turn to another
previous work and describe how to extend the framework there to the
setting with groups.
The approach to post-selection inference in \cite{lasso:fixed}
describes the selection event for the LASSO as a convex polytope
satisfying a list of affine constraints.
%stylized in Figure \ref{fig:polytope}.
If all groups are of size 1, then
the event that we observe a given set of variables
in the forward stepwise path (with their signs as they enter) would similarly
be given by a set of affine inequalities.
%This is described in more detail in \cite{exact:lars}.
After selection, exact inference for a (randomly) chosen contrast $\eta(y)^T\mu$ of the mean vector
$\mu$ could then be accomplished by analyzing the conditional distribution
\begin{equation}
y \sim N(\mu, \sigma^2 I) \bigl| Ay \leq b.
\end{equation}
See \cite{lasso:fixed,exact:lars} for further details on this approach.

With non-trivial groups, however, the event that $g^*(y)$, the first
group chosen, is equal to some fixed group $g$ is generally not given by
a set of affine constraints. Rather, considering line 3 of Algorithm~\ref{algo:fs}, we see
\begin{equation}
\label{eq:first:quadratic}
\begin{aligned}
\{g^*(y)=g\} &= \{ \|X_g^Ty\|_2 /w_g \geq \|X_h^Ty\|_2 /w_h , \ \forall h \neq g\} \\
&=  \{ y^T(X_g^TX_g)y / w_g^2 - y^T(X_h^TX_h)y / w_h^2 \geq 0, \ \forall h \neq g\} \\
\end{aligned}
\end{equation}
Hence, our selection event can be expressed as 
the intersection of a list of quadratic inequalities. This non-affine selection event is stylized in Figure
\ref{fig:curved}.
In the rest of this section, we consider an arbitrary selection procedure
$S$ that returns one of a set of possible outcomes $s \in {\cal S}$ by a set of quadratic inequalities.
That is, for all possible outcomes $s \in {\cal S}$ there is a set of indices
$I(s)$ such that
\begin{equation}
\label{eq:selection}
\begin{aligned}
\{S(y)=s\} = \cap_{i \in I(s)} \{y: y^TQ_iy + a_i^Ty\leq b_i \}.
\end{aligned}
\end{equation}
The quadratic forms are not assumed non-negative definite, but we can, without loss of generality
assume they are symmetric.
As in \cite{lasso:fixed} we will be interested in some randomly chosen contrast
$\eta(S(y))^T\mu$.
In our grouped selection procedure, the selection rule is $S(y)=g^*(y)$
and $\eta$ is typically 
$$
\eta(S(y)) = \eta(g^*(y)) = \frac{X_{g^*(y)}^Ty}{\|X_{g^*(y)}^Ty\|_2}.
$$

@@

%\subsection{Relation to covariance test}
%
%Consider the simple case for the first covariate to enter the model.
%In this case the covariance test statistic is $T = \lambda_1(\lambda_1 - M)$.
%\todo Does $M = \lambda_2$? Check proof, now that statistic has changed
%
%This is Lemma 5 from \cite{significance:lasso}.
%
%\begin{itemize}
%
%  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
%  \item Convergence to the limiting Exp(1) distribution is too slow
%    \[
%      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
%    \]
%    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
%  \item The limiting distribution only depends on $T$, but we also observe $M$
%  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
%  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
%    \[
%      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
%    \]
%    we can just use the left hand side
%  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
%  \item Now our p-value is
%    \[
%      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
%    \]   
%\end{itemize}
%
%

@@

We performed simulations with several classes of random design matrices
including Gaussian and categorical, and with fixed categorical designs
from the HIV data set of Section~\ref{sec:hiv}. Gaussian
design matrices are generated either independently or with some global
correlation $\rho > 0$ between all pairs of columns. Categorical matrices
were generated by first choosing a vector of probabilities for a given
variable from a Dirichlet prior, and then sampling category levels with
that vector of probabilities. Resampling was used to ensure the minimum
number of observations in any level was at least five. Finally, categorical
variables were encoded as groups of 0-1 vectors using the full encoding.

Since we are interested in variable selection we generate signals with
nonzero coefficients on the scale of $\gamma := \sqrt{2 \log(G)/n}$,
where recall $G$ is the number of groups.
Coefficients within a
nonzero group have roughly the same magnitude and are normalized
so that $\| \beta_g \|_2$ has the same scale independent of $p_g$.
The magnitudes for different
nonzero groups range from a lower limit times $\gamma$ to an upper
limit times $\gamma$. In text above each plot, the limits are listed
next to ``beta:'' and the numbers in parentheses are the largest
and smallest 2-norms of nonzero coefficient groups respectively.
Each plot also displays the
number of observations or rows, $n$, the number of columns, $p$, the
number of groups, $g$, and the largest and smallest group sizes in
parentheses if the groups are not all size 1. The number of nonzero
groups, $k$, is also displayed graphically by shading with the portion
of the plot showing steps after $k$ shaded gray.
The proportion of truly
nonzero groups recovered in the first $k$ steps is denoted $k$-TPP,
where recall $k$ is the number of truly nonzero groups.
